---
title: ''
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## s-worksheet - Extensions on the Likelihood Ratio Test

1. Snell's law \index{Snell's law} tell us how light bends at an interface - the angle of incidence versus the angle of refraction - based on the ratio of the velocities of light in the two isotropic media. If the angle of incidence of a laser beam in a vacuum is $\theta_1$ radians and the angle of refraction in an unknown medium is $\theta_2$, then 
$$n= \frac{\sin\theta_1}{\sin\theta_2}$$
is called the **index of refraction**  ($\beta=1/n$ is the velocity of light in the medium as a fraction of the speed of light in a vacuum.) Make repeated independent measurements in radians, $\theta_{1,1}, \theta_{1,2}, \cdots, \theta_{1,16}$ of the angle of incidence in vacuum and $\theta_{2,1}, \theta_{2,2}, \cdots, \theta_{2,16}$, of the angle of refraction in the second medium. If these measurements have 

* sample mean $\overline{\theta}_1=0.786$ radians and standard deviation 0.03 radians in vacuum and

* sample mean $\overline{\theta}_2=0. 326$ radians and standard deviation 0.06 radians in the unknown medium.

a. Snell's law gives an estimate $\hat n$ based on the values of  $\overline{\theta}_1$ and $\overline{\theta}_2$. Use the delta method to estimate the mean and standard deviation of $\hat n$.

\vfill 
b. You suspect that the substance is cubic zirconia ($n_z=2.165$) and not the claimed material, diamond ($n_d=2.418$). For the hypothesis
$$H_0: n=n_d\quad\mbox{versus}\quad H_1:n=n_z,$$
use the information above to devise a $z$-test for the hypothesis and report a $p$-value for the test.

```{r}
# First we need to find the z-value (Which is wrt H0) 
# using n = 2.209 and sd = 0.0994 which we found in Part (a)
# z = (n-n_d)/sd
z<-(2.209 - 2.418)/0.0994

# Now find the p-value using the pnorm function on z
(p<-pnorm(z))
```

\vfill 
c. Describe what the $p$-value communicates in this case.

In this case, the p-value tells us that we can reject the hypothesis at alpha = 0.02, or the 2% level. This means that we can be 98% certain that the sample we received is not actually diamond.

\vfill 